### 2024-06-16 Reports/Progress Report/sections/background.tex
added method section, paraphrased LLMs' background section

>>>>

\newpage
\subsection{Large Language Models}

The development of large language models (LLMs) represents a significant milestone in the field of artificial intelligence (AI) and natural language processing (NLP). The journey to modern LLMs has been marked by several key advancements in machine learning, computational power, and linguistic theory.

The earliest attempts at natural language processing can be traced back to the 1950s and 1960s, with efforts to create rule-based systems that could understand and generate human language. The most notable project from this era was the Georgetown-IBM experiment in 1954, which involved a rudimentary machine translation of Russian to English \cite{hutchins2004georgetown}. However, the complexity of human language posed significant challenges to these early systems.

The advent of machine learning in the 1980s introduced statistical methods for NLP, enabling the analysis of large corpora of text. One of the pioneering works was the development of hidden Markov models (HMMs) for speech recognition and part-of-speech tagging. These models utilized probabilities to predict sequences of words, marking a departure from purely rule-based systems \cite{rabiner1989tutorial}.

In the 2000s, the introduction of deep learning revolutionized NLP. Neural networks, particularly recurrent neural networks (RNNs) and long short-term memory (LSTM) networks, demonstrated the ability to capture long-range dependencies in text, which was crucial for tasks such as language modeling and machine translation. The development of word embeddings, such as Word2Vec, allowed for the representation of words in continuous vector spaces, preserving semantic relationships \cite{mikolov2013efficient}.

The transformer architecture, introduced by Vaswani et al. in 2017, marked a watershed moment in the development of LLMs. Transformers abstained from the sequential processing of RNNs in favor of a self-attention mechanism, which allowed for greater parallelization and the modeling of long-range dependencies more effectively \cite{vaswani2017attention}. This architecture underpinned models such as BERT (Bidirectional Encoder Representations from Transformers) and GPT (Generative Pre-trained Transformer).

BERT, introduced by Devlin et al. in 2018, leveraged bidirectional context, improving performance on a variety of NLP tasks through a two-stage process of pre-training and fine-tuning \cite{devlin2018bert}. GPT, initially introduced by Radford et al. in 2018 and significantly scaled up in subsequent versions, demonstrated the power of unsupervised pre-training on large text corpora followed by supervised fine-tuning \cite{Radford2018ImprovingLU}.

The latest iterations of GPT, particularly GPT-3 and beyond, exemplify the capabilities of LLMs with billions of parameters. These models are capable of generating human-like text, performing complex language tasks, and even exhibiting basic reasoning abilities \cite{brown2020language}. The success of these models is largely attributed to the scale of training data, advancements in computational resources, and sophisticated training techniques.

Overall, the development of LLMs has been driven by a confluence of advancements in statistical methods, neural network architectures, computational power, and the availability of large-scale datasets. These factors have collectively enabled the creation of models that can understand and generate human language with unprecedented proficiency.

\newpage
